### 2022-08-08 main/paper2.tex
Add files via upload

There have been many attempts during the past decades to capture and visualize the network of interconnections among a set of financial institutions. 
The most widely used concept of causality in time series econometrics is due to \citet{granger}. 
This is based on statistical analysis of the financial series such as their stock prices over a finite time period. 
Granger’s definition of causality states that a time series $X$ is a cause of another time series $Y$, if the mean squared error of the 1-step ahead forecast for $Y$ is smaller when the history of $X$ is included in the forecasting information set. 
Otherwise, when the forecast does not improve by including the information of $X$, then it is declared that $X$ does not cause $Y$.
This idea is reflected in the information-theoretic measure that we use in this work to infer the causal interactions among a network of time series.

In the great majority of practical applications, Granger-causality has been studied in the context of Vector Autoregressive (VAR) models. 
For instance, \citet{billio2012econometric} propose several measures  based on Granger-causality to capture the connections between the monthly returns of different financial institutions. 
It uses principle component analysis and pairwise Granger-causality tests to identify the causal networks. 
Other related works are \citet{diebold2014network} and \citet{barigozzi2016network} in which the authors propose connectedness measures based on generalized variance decomposition.
However, the measures introduced in these works are again limited to linear systems and they are based on pairwise comparison which as we show in Section  \ref{sec:dig} fails to infer the true causal relationships.
%and they are based on pairwise comparison which as we show in Section \ref{sec:dig}, it fail to infer the true causal relationships.
 %focuses on one particular network structure: the long-run variance decomposition network. Such networks are characterized by the infinite vector moving average and similar to \citet{diebold2014network}, they are limited to linear systems.

\clearpage
Contributions of this chapter are both in network identification literature and in finance.
%This paper adds contribution in the network identification literature as well as in finance literature. 
Our contributions to network identification are as follows:
\begin{itemize}
    \item We use an information-theoretic measure known as directed information (DI) to infer the Granger-causalities among a set of time series. 
This measure is non-parametric (i.e. it does not depend on the underlying model of the dynamics) and it is capable of capturing causal relationships in both linear and nonlinear systems. The output of this approach is a directed graph known as Directed Information Graph (DIG) that visualizes the interconnections among a set of time series such as stock returns.

\item Computing DI has high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
To overcome this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
\end{itemize}

@@

Applications of DIG are various in finance. 
In particular, we recommend to use it as a preliminary feature selection of another predictive model. Feature selection is a process often used in machine learning and statistics which consists of keeping only a subset of relevant features, usually to avoid overfitting or to reduce dimensionality. 
For example, \citet{fs1} and \citet{fs2} show that extraneous features are prone to reduce model's performance measures. 
Finance applications of feature selection models are various and include credit scoring, stock market behavior analysis  or even fraud detection (\citet{fs_fraud}).
\citet{tsai2009} states that feature selection preprocessing is not addressed carefully enough in the bankruptcy prediction literature. They compare five feature selection methods used in bankruptcy prediction: t-test, correlation matrix, step-wise regression, Principle Component Analysis (PCA) and factor analysis and show that any of these methods improves performance. 
\citet{fs_stock} uses a Sequential Forward Selection algorithm to select relevant features predicting the Turkish market index. The use of such feature selection model reduces model prediction error compared to the case where all features are used. This is due to information embedded in several economic factors already included in the market index.

@@

\subsection{Related Work}
In recent years, several approaches have been developed to generalize the applicability of Granger-causality to non-linear and large dynamics. 
To mention a few, \citet{psaradakis2005markov} introduce different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
In a similar context, \citet{bianchi2019modeling} investigate time-varying systemic risk based on a range of multi-factor asset pricing models and develop a Markov Chain Monte Carlo (MCMC) scheme to infer their model parameters and consequently obtain their corresponding networks.
Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger-causality.
%This work focuses on the linear factor model augmented with network dependence introduced by \citet{billio2018networks} and develops a method to identify causality among quantiles of the time series. 
This method is consistent with \citet{hong2009granger} and \citet{corsi2018measuring} that focus on causality among tail events. 
These methods are suitable for capturing causal relationships that are not in the center of their distributions, or in the mean but in the tails of their distributions.
It is important to emphasize that our proposed approach using DI is also capable of capturing such causal relations.

Most of the above aforementioned approaches are developed and tested for small size networks. 
Often, the problem of network identification in high dimensional settings requires more considerations and even its own techniques. 
For instance, \citet{billio2019bayesian} propose a Bayesian non-parametric Lasso prior (BNP-Lasso) for high-dimensional VAR models that can improve efficiency and accuracy. 
In order to overcome overparametrization and overfitting issues in large VAR models, BNP-Lasso clusters the VAR coefficients into groups and shrinks the coefficients of each group toward a common location.
%Although, the proposed method in this work is suitable for inference in high dimensional VAR models but 
However, this method is limited to linear models with Gaussian innovations.
To overcome this limitation, \citet{kalli2018bayesian} propose a Bayesian non-parametric approach that allows for nonlinearity in the conditional mean, heteroskedasticity in the conditional variance, and non-Gaussian innovations. However, unlike the BNP-Lasso, it does not allow sparsity in the model.
\citet{petrova2019quasi} proposes yet another non-parametric, quasi-Bayesian likelihood estimation methodology for high dimensional setting with time-varying parameters. 
%This method is also non-parametric. 
The work in \citet{iacopini2019bayesian} tackles the curse of dimensionality by a two-stage approach. In the first stage, a spike-and-slab prior distribution is used for each entry of the coefficient matrix which also identifies the interconnection network.
In the second stage, it imposes prior dependence on the coefficients by specifying a Markov process for their random distribution. 
%A side product of this procedure is the interconnection network.
A closely related work is \citet{bernardi2019high} that proposes a shrinkage and selection methodology designed for network inference in high-dimensional settings. It uses a regularized linear regression model with spike-and-slab prior on the parameters.
However, both methods are limited to VAR models.

The rest of the chapter is organized as follows. 
Section \ref{Sec1} reviews the notion of Granger-causality and formally introduces directed information graphs which are suitable for linear and nonlinear systems. 
In Section \ref{sec:method}, we introduce a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
Finally, in Section \ref{sec:exp}, we apply our method to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we use the daily stock prices of US firms.

@@

In this section, we present a statistical approach to learn the causal interconnections in a dynamical systems based on Granger-causality (see \citet{granger}). 
We begin by introducing some notations.
Plain capital letters denote random variables or processes, while lowercase letters denote their realizations. 
Bold letters are used for column vectors, matrices, and tensors and calligraphy letters are used for sets. 
We use $X_{j,t}$ to denote the value of a time series $X_j$ at time $t$ 
%and $X_{j,t_1}^{t_2}$, when $t_1\leq t_2$, to denote $X_j$ from time $t_1$ to $t_2$, namely, $(X_{j,t_1},...,X_{j,t_2})$. 
%Subsequently, 
and $X^{t}_j$ to denote the time series $X_j$ up to time $t$. 
For a set $\mathcal{A}=\{a_1,...,a_n\}$ and an index set $\mathcal{I}\subseteq\{1,...,n\}$, we define $\mathcal{A}_{-\mathcal{I}}:=\mathcal{A}\setminus\{a_i: i\in\mathcal{I}\}$.

@@

Researchers from different fields have developed various frameworks and graphical models to capture and represent interconnections among variables or processes.
One of the most popular and widely used frameworks in economics is the notion of Granger-causality.
The basic idea in this framework was originally introduced by \citet{wiener1956theory}, and later formalized by \citet{granger}. 
The idea is as follows: ``we say that $X$ is causing $Y$ if we are better able to predict the future of $Y$ using all available information than if the information apart from the past of $X$ had been used."

Despite broad philosophical viewpoint of \citet{granger1963economic}, his formulation for practical implementation was done using multivariate autoregressive (MVAR) models and linear regression which has been widely adopted in econometric and other disciplines.
More precisely, in order to identify the influence of $X_t$ on $Y_t$ in a MVAR comprises of three time series $\{X,Y,Z\}$, Granger's idea is to compare the performance of two linear regressions: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$. Clearly, the performance of the second predictor is bounded by the first predictor. If they have the same performance, then we say $X$ does not Granger-cause $Y$. It is important to emphasize that this formulation is only applicable in linear systems.

To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger-causality. 
In this work, we introduce and apply directed information (see \citet{directed}), an information-theoretical tools to measure interconnections among firms.
Directed information (DI) has been used in many applications to infer causal relationships. For example, \citet{spike} and \citet{di12} use it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} apply to market data.

@@

In the rest of this section, we describe how the DI can capture the interconnections in causal dynamical systems (linear or non-linear) and formally define DIGs.

Consider a dynamical system comprised of three time series $\{X,Y,Z\}$ that we assume they have a joint probability density function $p(X,Y,Z)$. 
To answer whether $X$ has influence on $Y$ or not over time horizon $[1,T]$,  following the idea of Granger, we compare the average performance of two particular predictors over this time horizon. 
The first predictor uses the history of all three time series while the second one uses the history of all processes excluding process $X$. 
On average, the performance of the predictor with less information (the second one) is upper bounded by the performance of the predictor with more information  (the first one). 
However,  if the prediction of both predictors are close over time horizon $[1,T]$,  it is an indication that $X$ does not cause $Y$ in this time horizon.
To rigorously formalize this idea, we need the predictors and a measure to compare their performances.

@@

Given the predictions of the first and the second predictors at time $t$ for an outcome $y_t\in \mathcal{Y}$, the goodness of these predictions are measured by the log-loss that are defined respectively by 
\begin{align*}
    &-\log p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1}),\\
    &-\log p(Y_t=y_t|Y^{t-1},Z^{t-1}).
\end{align*}
According to the above measures of goodness, the better the predictor is, the smaller its log-loss will be.
This loss function has meaningful information-theoretical interpretations.
Namely, the log-loss is the Shannon's code length\footnote{It is also called the description length of $y_t$. For more information see \citet{cover2012elements}.}, i.e., the number of bits required to efficiently represent $y_t$.

At time $t$ for an outcome $y_t\in \mathcal{Y}$, the difference between the log-losses of the two predictors compares their performances.  This difference is also called regret, denoted $r_t$ :
\begin{align}
r_t:&=-\log{p(Y_t=y_t|Y^{t-1},Z^{t-1})}-\big(-\log {p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}\big) \nonumber \\
&=\log \frac{p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}{p(Y_t=y_t|Y^{t-1},Z^{t-1})}.
\end{align}
Note that the regrets are non-negative for all $t$ and all outcomes $y_t$.
The average regret over the time horizon $[1,T]$ is given by 
\begin{align}\label{eq:ave}
   \frac{1}{T}\sum_{t=1}^T \mathbb{E}[r_t],
\end{align}
where the expectation is taken over the joint density function\footnote{For the sake of notational simplicity, we use $p(y^{t},z^{t-1},x^{t-1})$ to denote $p(Y^{t}=y^{t},Z^{t-1}=z^{t-1},X^{t-1}=x^{t-1})$.} of $X$, $Y$, and $Z$, i.e., 
\begin{align}
    \mathbb{E}[r_t]=\int p(y^{t},z^{t-1},x^{t-1})\log \frac{p(y_t|y^{t-1},z^{t-1},x^{t-1})}{p(y_t|y^{t-1},z^{t-1})}dy^tdx^{t-1}dz^{t-1}.
\end{align}
The average regret in \eqref{eq:ave} is called directed information (DI) and will be our measure of causation in this work.
This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$. 
This definition can be generalized to more than three time series as follows,

@@

It is important to emphasize that the definition of DI does not rely on any model assumption, thus DI is capable of inferring the causal relationships in general (linear or non-linear) dynamical systems. Next, we define the graphical model that we use in this work to visualize the causal network among firms.

@@

A simple way to represent the DIG $G$ of a dynamical system is via the adjacency matrix $\textbf{DIG}=[d_{i,j}]_{m\times m}$ that is defined by
\begin{align}
    d_{j,i} = \left\{
\begin{array}{ll}
      1 & \text{if}\  I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})>0,\\
      0 & \text{otherwise}. \\
\end{array} 
\right. 
\end{align}
Given a DIG $G=(\mathcal{V},\mathcal{E})$, we define the parent set of node $R_j$ denoted by $\mathcal{PA}_j\subset \mathcal{V}$ to be the set of all times series that have direct influences on $R_j$, i.e., $\mathcal{PA}_j:=\{R_k: d_{j,k}=1\}$. Similarly, the children set of node $R_j$ is given by $\mathcal{CH}_j:=\{R_k: d_{k,j}=1\}$. 
Example \ref{r1} demonstrates the DIG of a simple linear system.
\begin{example}\label{r1}
Consider a network of three times series $\{X,Y,Z\}$ with the following linear dynamic,
\begin{small}
\begin{equation}\label{model1}
\begin{pmatrix}
X_t \\
Y_t\\
Z_t
\end{pmatrix}=\begin{pmatrix}
0.5 & 0& 0\\
0.4 & 0.5 & 0\\
0 & -0.2 & 0
\end{pmatrix}\begin{pmatrix}
X_{t-1} \\
Y_{t-1}\\
Z_{t-1}
\end{pmatrix}+\begin{pmatrix}
N_{X_t} \\
N_{Y_t}\\
N_{Z_t}
\end{pmatrix},
\end{equation}
\end{small}
where $N_X$, $N_Y$, and $N_Z$ are three independent stationary Gaussian processes with zero mean and a diagonal covariance matrix (1, 0.9, 1). 
Since the dynamic is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in more general settings.}(see \citet{acc2014}).
\begin{equation}\label{simueq}
I(Z\rightarrow Y||X)=\frac{1}{2T}\sum_{t=1}^{T}\log\frac{|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},X_{t-1}}|}{|\Sigma_{Y_{t-1},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},Y_{t},X_{t-1}}|},
\end{equation}
where $|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}|$ denotes the determinant of the covariance matrix of $\{Y_{t-1},Y_{t},X_{t-1}\}$. Using \eqref{simueq}, we compute the DIs of this system,
\begin{small}
\begin{align}  \label{di_xyz}
&I(Y\rightarrow X||Z)=0,\ \ I(Z\rightarrow X||Y)=0,\ \ I(Z\rightarrow Y||X)=0,\ \ I(Z\rightarrow Z||X,Y)=0,\\
&I(X\rightarrow Z||Y)=0,\ \ \ \ I(X\rightarrow Y||Z)\approx 0.1, \ \ \ I(Y\rightarrow Z||X)\approx 0.03. \nonumber
\end{align}
\end{small}
Figure \ref{fig:example_dig} illustrates the DIG of the system \eqref{model1}. In this example, $\mathcal{PA}_Z=\{Y\}$ and $\mathcal{CH}_Z=\{\}$.
\begin{figure}[h]
\hspace{4cm}
\xygraph{ !{<0cm,0cm>;<1.9cm,0cm>:<0cm,1.7cm>::} 
!{(-.7,1) }*+{X}="x"
!{(.4,1.6) }*+{Y}="y"
!{(1.5,1)}*+{Z}="z"
%"x":@(ld,lu)"x"
%"y":@(lu,ru)"y"
"x":"y"
"y":"z"
}
\mycaption{DIG estimated from the directed informations \eqref{di_xyz}}{X influences Y, Y influences Z. This DIG correctly estimates the dynamics of the system in \eqref{model1}.}
\label{fig:example_dig}
\end{figure}
\end{example}

@@

\begin{remark}
A causal model allows a factorization of the joint density function in some specific ways.
It was shown in \citet{directed} that under a mild assumption, the joint density function of a causal discrete-time dynamical system with DIG $G(\mathcal{R},\mathcal{E})$ can be factorized as follows,
\begin{equation}\label{abo}
p(R_1,...,R_m)=\prod_{i=1}^{m}\prod_{t=1}^T p(R_{i,t}|\mathcal{R}_{\mathcal{PA}_i\cup\{i\}}^{t-1}).
\end{equation}
%where $\mathcal{B}_i$ denotes the parent set of node $R_i$ in graph $G$ and $\mathcal{R}_{B_i\cup\{i\}}^{t-1}$ denotes the time series with indices $B_i\cup\{i\}$ up to time $t-1$.  
Such factorization is called a generative model. 
\end{remark}

>>>>

There have been many attempts during the past decades to capture and visualize the network of interconnections among a set of financial institutions. 
A well-known time series causality measure used in econometrics is Granger-causality (\citet{granger}).
This is based on statistical analysis of the financial time series such as their stock prices over a finite time period. 
Granger’s definition of causality states that a time serie $X$ is a cause of another time serie $Y$, if the one-step future forecast of $Y$ is more precise when the forecasting information set includes X. 
Otherwise, when the forecast does not improve by including the information of $X$, then it is declared that $X$ does not cause $Y$.
This idea is reflected in the information-theoretic measure that we use in this work to infer the causal interactions among a network of time series.

Most empirical applications of Granger-causality have been studied with Vector Autoregressive (VAR) models. 
For instance, \citet{billio2012econometric} propose several measures based on Granger-causality to capture the interconnections between the returns of financial institutions on a monthly basis. 
It uses principle component analysis and pairwise Granger-causality tests to identify the causal networks. 
Other related works are \citet{diebold2014network} and \citet{barigozzi2016network} in which the authors propose connectedness measures based on generalized variance decomposition.
However, the measures introduced in these works are again limited to linear systems and they are based on pairwise comparison which as we show in Section  \ref{sec:dig} fails to infer the true causal relationships.
%and they are based on pairwise comparison which as we show in Section \ref{sec:dig}, it fail to infer the true causal relationships.
 %focuses on one particular network structure: the long-run variance decomposition network. Such networks are characterized by the infinite vector moving average and similar to \citet{diebold2014network}, they are limited to linear systems.

\clearpage
Contributions of this chapter are both in network identification literature and in finance.
%This paper adds contribution in the network identification literature as well as in finance literature. 
Our contributions to network identification are as follows:
\begin{itemize}
    \item We use directed information (DI), an information measure to infer the Granger-causalities among a set of time series. 
This measure is non-parametric (i.e. it does not depend on the underlying model of the dynamics) and it is capable of capturing causal relationships in both linear and nonlinear systems. The output of this approach is a directed graph known as Directed Information Graph (DIG) that visualizes the interconnections among a set of time series such as stock returns.

\item Computing DI has high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
To solve this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
\end{itemize}

@@

Applications of DIG are various in finance. 
In particular, we recommend to use it as a preliminary feature selection of another predictive model. Feature selection is a process often used in machine learning and statistics which consists of keeping only a subset of relevant features, usually to avoid overfitting or to reduce dimensionality. 
For example, \citet{fs1} and \citet{fs2} show that extraneous features are prone to reduce model's performance measures. 
Finance applications of feature selection models are various and include credit scoring, stock market behavior analysis  or even fraud detection (\citet{fs_fraud}).
\citet{tsai2009} states that feature selection preprocessing is not addressed carefully enough in the bankruptcy prediction literature. They compare five feature selection methods used in bankruptcy prediction: step-wise regressions, correlation matrix, principal component analysis, t-tests and factor analysis and show that any of these methods improves performance. 
\citet{fs_stock} uses a Sequential Forward Selection algorithm to select relevant features predicting the Turkish market index. The use of such feature selection model reduces model prediction error compared to the case where all features are used. This is due to information embedded in several economic factors already included in the market index.

@@

\subsection{Related Work}
In recent years, several approaches have been developed to generalize the applicability of Granger-causality to non-linear and large dynamics. 
To mention a few, \citet{psaradakis2005markov} introduce different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
In a similar context, \citet{bianchi2019modeling} investigate time-varying systemic risk based on a range of multi-factor asset pricing models and develop a Markov Chain Monte Carlo scheme to infer their model parameters and consequently obtain their corresponding networks.
Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger-causality.
%This work focuses on the linear factor model augmented with network dependence introduced by \citet{billio2018networks} and develops a method to identify causality among quantiles of the time series. 
This method is consistent with \citet{hong2009granger} and \citet{corsi2018measuring} that focus on causality among tail events. 
These methods are suitable for capturing causal relationships that are not in the center of their distributions, or in the mean but in the tails of their distributions.
It is important to emphasize that our proposed approach using DI is also capable of capturing such causal relations.

Most of the above aforementioned approaches are developed and tested for small size networks. 
Often, the problem of network identification in high dimensional settings requires more considerations and even its own techniques. 
For instance, \citet{billio2019bayesian} propose a Bayesian non-parametric Lasso (BNPL) prior for high-dimensional vector autoregressive models that improve efficiency and accuracy. 
To overcome overfitting in large VAR models, BNPL clusters the vector autoregressive coefficients and shrinks the coefficients of each cluster toward a common location.
%Although, the proposed method in this work is suitable for inference in high dimensional VAR models but 
However, this method is limited to linear models with Gaussian innovations.
To overcome this limitation, \citet{kalli2018bayesian} propose a Bayesian non-parametric approach that allows for nonlinearity in the conditional mean, heteroskedasticity in the conditional variance, and non-Gaussian innovations. However, unlike the BNP-Lasso, it does not allow sparsity in the model.
\citet{petrova2019quasi} proposes yet another non-parametric, quasi-Bayesian likelihood estimation methodology for high dimensional setting with time-varying parameters. 
%This method is also non-parametric. 
The work in \citet{iacopini2019bayesian} tackles the curse of dimensionality by a two-stage approach. First, a spike-and-slab prior distribution is used for each entry of the coefficient matrix which also identifies the interconnection network.
In the second stage, it imposes prior dependence on the coefficients by specifying a Markov process for their random distribution. 
%A side product of this procedure is the interconnection network.
A closely related work is \citet{bernardi2019high} that proposes a shrinkage and selection methodology designed for network inference in high-dimensional settings. It uses a regularized linear regression model with spike-and-slab prior on the parameters.
However, both methods are limited to VAR models.

The remainder of this chapter is strutured as follows. 
Section \ref{Sec1} reviews the notion of Granger-causality and formally introduces directed information graphs which are suitable for linear and nonlinear systems. 
Section \ref{sec:method} introduces a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
Section \ref{sec:exp} applies our methodology to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we use the daily stock prices of US firms.

@@

We present in this section our econometric approach to learn the causal dependencies in a dynamical systems based on Granger-causality (see \citet{granger}). 
We begin by introducing some notations.
Plain capital letters denote random variables or processes, while lowercase letters denote their realizations. 
Bold letters are used for column vectors, matrices, and tensors and calligraphy letters are used for sets. 
We use $X_{j,t}$ to denote the value of a time series $X_j$ at time $t$ 
%and $X_{j,t_1}^{t_2}$, when $t_1\leq t_2$, to denote $X_j$ from time $t_1$ to $t_2$, namely, $(X_{j,t_1},...,X_{j,t_2})$. 
%Subsequently, 
and $X^{t}_j$ to denote the time series $X_j$ up to time $t$. 
For a set $\mathcal{A}=\{a_1,...,a_n\}$ and an index set $\mathcal{I}\subseteq\{1,...,n\}$, we define $\mathcal{A}_{-\mathcal{I}}:=\mathcal{A}\setminus\{a_i: i\in\mathcal{I}\}$.

@@

Various frameworks and graphical models have been developed to capture and represent interconnections among variables or processes.
One of the most popular and widely used frameworks in economics is the notion of Granger-causality originally introduced by \citet{wiener1956theory} and
generalized by \citet{granger}. We say that $X$ Granger-causes $Y$ if the prediction of the future of $Y$ is more precise using all available information (including $X$) than using all available information except $X$.

This formulation was originally implemented using multivariate autoregressive (MVAR) models and linear regressions.
In particular, let $\{X,Y,Z\}$ be three time series. In order to identify the influence of $X_t$ on $Y_t$, Granger's idea is to compare the performance of two linear regressions\footnote{Note that this formulation is only applicable in linear systems.}: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$. If they have the same performance, then we say $X$ does not Granger-cause $Y$.

To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger-causality. 
In this work, we introduce and apply directed information (see \citet{directed}), an information-theoretical tool to measure interconnections among firms.
Directed information (DI) has been used in many applications to infer causal relationships. For example, \citet{spike} and \citet{di12} use it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} apply it to market data.

@@

We describe in this section how the directed information is able to capture the connections in causal networks. Next, we formally define directed information graphs (DIG).

We define a dynamical system constituted of three time series $\{X,Y,Z\}$ that we assume have a joint probability density function $p(X,Y,Z)$. 
Granger-causality states that to know whether $X$ influences $Y$ or not, we need to compare the performance of two predictors of $Y$.
The first predictor uses the history of all information available (i.e. $\{X,Y,Z\}$) while the second predictor uses only the history of $\{Y,Z\}$
If the former performs better than the latter, $X$ has information on $Y$. However, if they perform equally, it is an indication that $X$ is not causing $Y$ in this time interval.
To rigorously formalize this idea, we need the predictors and a measure to compare their performances.

@@

Given the predictions of the first and the second predictors at time $t$ for an outcome $y_t\in \mathcal{Y}$, the goodness of these predictions are measured by the log-loss that are defined respectively by 
\begin{align*}
    &-\log p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1}),\\
    &-\log p(Y_t=y_t|Y^{t-1},Z^{t-1}).
\end{align*}
According to the above measures of goodness, the better the predictor is, the smaller its log-loss will be.
This loss function also has meaningful information-theoretical interpretations.
Namely, the log-loss is the Shannon's code length\footnote{It is also called the description length of $y_t$. For more information see \citet{cover2012elements}.}, i.e., the number of bits required to efficiently represent $y_t$ (see \citet{etesami2018econometric}).

At time $t$ for an outcome $y_t\in \mathcal{Y}$, the difference between the log-losses of the two predictors compares their performances.  This difference is also called regret, denoted $r_t$ :
\begin{align}
r_t:&=-\log{p(Y_t=y_t|Y^{t-1},Z^{t-1})}-\big(-\log {p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}\big) \nonumber \\
&=\log \frac{p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}{p(Y_t=y_t|Y^{t-1},Z^{t-1})}.
\end{align}
Regrets are always positive for all $t$ and all outcomes $y_t$.
Over the time interval $[1,T]$, the average regret is
\begin{align}\label{eq:ave}
   \frac{1}{T}\sum_{t=1}^T \mathbb{E}[r_t],
\end{align}
where the expectation is taken over the joint density function\footnote{For the sake of notational simplicity, we use $p(y^{t},z^{t-1},x^{t-1})$ to denote $p(Y^{t}=y^{t},Z^{t-1}=z^{t-1},X^{t-1}=x^{t-1})$.} of $X$, $Y$, and $Z$, i.e., 
\begin{align}
    \mathbb{E}[r_t]=\int p(y^{t},z^{t-1},x^{t-1})\log \frac{p(y_t|y^{t-1},z^{t-1},x^{t-1})}{p(y_t|y^{t-1},z^{t-1})}dy^tdx^{t-1}dz^{t-1}.
\end{align}
The average regret in \eqref{eq:ave} is the directed information (DI). We use it as the measure of causation in this chapter.
This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$. 
This definition can be generalized to more than three time series as follows,

@@

Note that the definition of DI does not rely on any model assumption, thus DI is capable of inferring the causal relationships in general (linear or non-linear) dynamical systems. Next, we define the graphical model that we use in this work to visualize the causal network among firms.

@@

A simple way to represent the DIG $G$ of a dynamical system is via the adjacency matrix $\textbf{DIG}=[d_{i,j}]_{m\times m}$ that is defined by
\begin{align}
    d_{j,i} = \left\{
\begin{array}{ll}
      1 & \text{if}\  I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})>0,\\
      0 & \text{otherwise}. \\
\end{array} 
\right. 
\end{align}
Given a DIG $G=(\mathcal{V},\mathcal{E})$, we define the parent set of node $R_j$ denoted by $\mathcal{PA}_j\subset \mathcal{V}$ to be the set of all times series that have direct influences on $R_j$, i.e., $\mathcal{PA}_j:=\{R_k: d_{j,k}=1\}$. Similarly, the children set of node $R_j$ is given by $\mathcal{CH}_j:=\{R_k: d_{k,j}=1\}$. 
Example \ref{r1} demonstrates the DIG of a simple linear system.

\begin{example}\label{r1}
Let $\{X,Y,Z\}$ be a network of three time series with the following dynamics,
\begin{small}
\begin{equation}\label{model1}
\begin{pmatrix}
X_t \\
Y_t\\
Z_t
\end{pmatrix}=\begin{pmatrix}
0.5 & 0& 0\\
0.4 & 0.5 & 0\\
0 & -0.2 & 0
\end{pmatrix}\begin{pmatrix}
X_{t-1} \\
Y_{t-1}\\
Z_{t-1}
\end{pmatrix}+\begin{pmatrix}
N_{X_t} \\
N_{Y_t}\\
N_{Z_t}
\end{pmatrix},
\end{equation}
\end{small}
where $N_X$, $N_Y$, and $N_Z$ are three independent stationary Gaussian processes with zero mean and a diagonal covariance matrix (1, 0.9, 1). 
Since the dynamics is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in more general settings.}(see \citet{acc2014}).
\begin{equation}\label{simueq}
I(Z\rightarrow Y||X)=\frac{1}{2T}\sum_{t=1}^{T}\log\frac{|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},X_{t-1}}|}{|\Sigma_{Y_{t-1},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},Y_{t},X_{t-1}}|},
\end{equation}
where $|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}|$ denotes the determinant of the covariance matrix of $\{Y_{t-1},Y_{t},X_{t-1}\}$. Using \eqref{simueq}, we compute the DIs of this system,
\begin{small}
\begin{align}  \label{di_xyz}
&I(Y\rightarrow X||Z)=0,\ \ I(Z\rightarrow X||Y)=0,\ \ I(Z\rightarrow Y||X)=0,\ \ I(Z\rightarrow Z||X,Y)=0,\\
&I(X\rightarrow Z||Y)=0,\ \ \ \ I(X\rightarrow Y||Z)\approx 0.1, \ \ \ I(Y\rightarrow Z||X)\approx 0.03. \nonumber
\end{align}
\end{small}
Figure \ref{fig:example_dig} illustrates the DIG of the system \eqref{model1}. In this example, $\mathcal{PA}_Z=\{Y\}$ and $\mathcal{CH}_Z=\{\}$.
\begin{figure}[h]
\hspace{4cm}
\xygraph{ !{<0cm,0cm>;<1.9cm,0cm>:<0cm,1.7cm>::} 
!{(-.7,1) }*+{X}="x"
!{(.4,1.6) }*+{Y}="y"
!{(1.5,1)}*+{Z}="z"
%"x":@(ld,lu)"x"
%"y":@(lu,ru)"y"
"x":"y"
"y":"z"
}
\mycaption{DIG estimated from the directed informations \eqref{di_xyz}}{X influences Y, Y influences Z. This DIG correctly estimates the dynamics of the system in \eqref{model1}.}
\label{fig:example_dig}
\end{figure}
\end{example}

@@

\begin{comment}
\begin{remark}
A causal model allows a factorization of the joint density function in some specific ways.
It was shown in \citet{directed} that under a mild assumption, the joint density function of a causal discrete-time dynamical system with DIG $G(\mathcal{R},\mathcal{E})$ can be factorized as follows,
\begin{equation}\label{abo}
p(R_1,...,R_m)=\prod_{i=1}^{m}\prod_{t=1}^T p(R_{i,t}|\mathcal{R}_{\mathcal{PA}_i\cup\{i\}}^{t-1}).
\end{equation}
%where $\mathcal{B}_i$ denotes the parent set of node $R_i$ in graph $G$ and $\mathcal{R}_{B_i\cup\{i\}}^{t-1}$ denotes the time series with indices $B_i\cup\{i\}$ up to time $t-1$.  
Such factorization is called a generative model. 
\end{remark}
\end{comment}
